---
title: "The Basics of Networks"
author: "James D. Wilson"
date: "7/2/2017"
output: html_document
---

##Important Graph Analytic Packages

Two important packages that are commonly used for network analysis include `igraph` and `statnet`. Full documentation on `igraph` is available at http://igraph.org/redirect.html, including functions in the package for \texttt{Python}, \texttt{R} and \texttt{C}. 

Let's begin by installing these packages.

```{r, echo = TRUE, eval = TRUE}
install.packages("statnet", repos = "https://cloud.r-project.org")
install.packages("igraph", repos = "https://cloud.r-project.org")
install.packages("Matrix", repos = "https://cloud.r-project.org")
library(igraph)
library(statnet)
library(Matrix)

```

## EXAMPLE 1: Zachary's Karate Club Network
### Download Data and Convert to Matrix
```{r, echo = TRUE, eval = TRUE}
#download raw edgelist
karate.data <- read.table("https://raw.githubusercontent.com/jdwilson4/Network-Analysis-I/master/Data/karate.txt", sep = " ", header = TRUE, stringsAsFactors = FALSE)
#convert data to matrix
karate.edgelist <- matrix(unlist(karate.data), ncol = 2) + 1
```

### Convert Network Edge Data to 'igraph' Objects
```{r, echo = TRUE, eval = TRUE}
#general igraph (defaults to directed)
karate.igraph <- graph.edgelist(karate.edgelist)
#undirected igraph
karate.undirected.graph <- graph.edgelist(karate.edgelist, directed = FALSE)
#adjacency matrix from igraph object
karate.adjacency <- as_adj(karate.igraph)

```

### Adjacency Matrix Visualization
```{r, echo = TRUE, eval = TRUE}
image(Matrix(karate.adjacency))

```

### Create network from edge list
Try running this block a few times. What happens to the network?
```{r, echo = TRUE, eval = TRUE}
karate.network <- network(karate.edgelist)
plot(karate.network, main = paste("Zachary's Karate Club"), usearrows = TRUE, edge.col = "grey50")

```
```{r, echo = TRUE, eval = TRUE}
karate.x <- plot(karate.network, main = paste("Zachary's Karate Club"), usearrows = TRUE, edge.col = "grey50")
```

###Color Vertices According to Known Labels
```{r, echo = TRUE, eval = TRUE}
karate.labels <- c(1,1,1,1,1,1,1,1,1,2,1,1,1,1,2,2,1,1,2,1,2,1,2,2,2,2,2,2,2,2,2,2,2,2)
plot(karate.network, main = "Zachary's Karate Network", usearrows = TRUE, edge.col = "grey50", coord = karate.x, vertex.col = karate.labels)


```

## EXAMPLE 2: Political Blogs Network
### Download Data and Convert to Matrix
```{r, echo = TRUE, eval = TRUE}
#download raw edgelist
pblog.data <- read.table("https://raw.githubusercontent.com/jdwilson4/Network-Analysis-I/master/Data/polblogs.txt", sep = " ", header = TRUE, stringsAsFactors = FALSE)
#convert data to matrix
pblog.edgelist <- as.matrix(pblog.data) + 1
```

### Convert Network Edge Data to 'igraph' Objects
```{r, echo = TRUE, eval = TRUE}
#general igraph (defaults to directed)
pblog.igraph <- graph.edgelist(pblog.edgelist)
#undirected igraph
pblog.undirected.graph <- graph.edgelist(pblog.edgelist, directed = FALSE)
#adjacency matrix from igraph object
pblog.adjacency <- as_adj(pblog.igraph)

```

### Adjacency Matrix Visualization
```{r, echo = TRUE, eval = TRUE}
image(Matrix(pblog.adjacency))

```

###Create network from edge list
```{r, echo = TRUE, eval = TRUE}
pblog.network <- network(pblog.edgelist)
pblog.x <- plot(pblog.network, main = paste("Political Blog Network"), usearrows = TRUE, edge.col = "grey50", vertex.col = "yellow")

```

###Color Vertices According to Known Labels & Plot
```{r, echo = TRUE, eval = TRUE}
# original labels: 
# [1:758] = 0 (left or liberal), blue
# [759: 1490] = 1 (right or conservative), red
# color labels : 0 --> 4 = blue , 1 --> 2 = red
pblog.labels <- append(rep(4, 758), rep(2, 732), after = 758)
plot(pblog.network, main = "Political Blog Network", usearrows = TRUE, edge.col = "grey50", coord = pblog.x, vertex.col = pblog.labels)
```
