---
title: "Identifying the Most Central Folks in a Network"
author: "James D. Wilson and Melanie Baybay"
date: "7/11/2017"
output: html_document
---


## Network Analysis Setup

Two important packages that are commonly used for network analysis include `igraph` and `statnet`. Full documentation on `igraph` is available at http://igraph.org/redirect.html, including functions in the package for \texttt{Python}, \texttt{R} and \texttt{C}. 

Since we've previously installed these packages, let's begin by loading them into our library.

```{r, echo = TRUE, eval = TRUE}
library(igraph)
library(statnet)
library(Matrix)

```

## EXAMPLE 1: Zachary's Karate Club Network
### Download Data and Convert to Matrix
```{r, echo = TRUE, eval = TRUE}
#download raw edgelist
karate.data <- read.table("https://raw.githubusercontent.com/jdwilson4/Network-Analysis-I/master/Data/karate.txt", sep = " ", header = TRUE, stringsAsFactors = FALSE)
#convert data to matrix
karate.edgelist <- matrix(unlist(karate.data), ncol = 2) + 1
```

### Convert Network Edge Data to 'igraph' Objects
```{r, echo = TRUE, eval = TRUE}
#general igraph (defaults to directed)
karate.igraph <- graph.edgelist(karate.edgelist)
#undirected igraph
karate.undirected.graph <- graph.edgelist(karate.edgelist, directed = FALSE)
#adjacency matrix from igraph object
karate.adjacency <- as_adj(karate.igraph)

```

### Adjacency Matrix Visualization
```{r, echo = TRUE, eval = TRUE}
image(Matrix(karate.adjacency))

```

---

##Network Centrality Measures and Visualization
###Create network from edge list
```{r, echo = TRUE, eval = TRUE}
karate.network <- network(karate.edgelist)
# known labels
group.labels <- c(1,1,1,1,1,1,1,1,1,2,1,1,1,1,2,2,1,1,2,1,2,1,2,2,2,2,2,2,2,2,2,2,2,2) 
plot(karate.network, main = paste("Zachary's Karate Club"), usearrows = TRUE, edge.col = "grey50", vertex.col = group.labels)

```
```{r, echo = TRUE, eval = TRUE}
#store the node and edge coordinates
x <- plot(karate.network, main = paste("Zachary's Karate Club"), usearrows = TRUE, edge.col = "grey50", vertex.col = group.labels)

#change the size of each vertex according to out-degree
plot(karate.network, main = paste("Out-Degree Centrality"), usearrows = TRUE, vertex.cex = rowSums(as.matrix(karate.adjacency)) + 1, edge.col = "grey50", coord = x, vertex.col = group.labels)

#change the size of each vertex according to in-degree
plot(karate.network, main = paste("In-Degree Centrality"), usearrows = TRUE, vertex.cex = colSums(as.matrix(karate.adjacency)) + 1, edge.col = "grey50", coord = x, vertex.col = group.labels)
```

###Calculate In-Degree, Out-Degree, Eigenvector, Betweenness, and Closeness Centralities
- **in-degree centrality:** number of neighbors that point to a node in a directed network
- **out-degree centrality:** number of neighbors that a node points to in a directed network
- **eigenvector centrality:** measures the importance of node based on how influential its neighbors are to the network
```{r, echo = TRUE, eval = TRUE}
#in degree centrality
in.degree.centrality <- colSums(as.matrix(karate.adjacency)) + 1

#out degree centrality
out.degree.centrality <- rowSums(as.matrix(karate.adjacency)) + 1

#eigenvector centrality
eigenvector.centrality <- eigen_centrality(karate.igraph, directed = FALSE)$vector
```
- **betweenness centrality:** based on shortest paths, it represents the degree to which nodes lie between one another 
- **closeness centrality: ** average length of the shortest paths from a node to all other nodes in the network


First, let's identify what the shortest paths are between each pair of nodes. The following treats shortest path calculations as the shortest path of the corresponding undirected graph.
```{r, echo = TRUE, eval = TRUE}
shortest.distances <- distances(karate.igraph, mode = "all") 

#network diameter: the longest shortest path that is not infinite
max(shortest.distances[which(shortest.distances < Inf)])

#normalized betweenness centrality
betweenness.centrality <- estimate_betweenness(karate.igraph, directed = TRUE, cutoff = 10)

#closeness centrality
closeness.centrality <- estimate_closeness(karate.igraph, mode = "total", normalized = FALSE, cutoff = 10)
```

###Plot Network According to Centrality Measures
```{r, echo = TRUE}
par(mfrow = c(2,3))

#plot original network 
plot(karate.network, main = "Zachary's Karate Network", usearrows = TRUE, edge.col = "grey50", coord = x, vertex.col = group.labels)

#plot in-degree
plot(karate.network, main = paste("In-Degree Centrality"), usearrows = TRUE, vertex.cex = in.degree.centrality, edge.col = "grey50", coord = x, vertex.col = group.labels)

#plot out-degree
plot(karate.network, main = paste("Out-Degree Centrality"), usearrows = TRUE, vertex.cex = out.degree.centrality, edge.col = "grey50", coord = x, vertex.col = group.labels)

#plot eigenvector
plot(karate.network, main = paste("Eigenvector Centrality"), usearrows = TRUE, vertex.cex = eigenvector.centrality*4 + 1, edge.col = "grey50", coord = x, vertex.col = group.labels)

#plot betweenness
plot(karate.network, main = paste("Betweenness Centrality"), usearrows = TRUE, vertex.cex = betweenness.centrality / 2 + 1, edge.col = "grey50", coord = x, vertex.col = group.labels)

#plot closeness
plot(karate.network, main = paste("Closeness Centrality"), usearrows = TRUE, vertex.cex = closeness.centrality * 150 + 1, edge.col = "grey50", coord = x, vertex.col = group.labels)


```

## EXAMPLE 2: Political Blogs Network
### Download Data and Convert to Matrix
```{r, echo = TRUE, eval = TRUE}
#download raw edgelist
pblog.data <- read.table("https://raw.githubusercontent.com/jdwilson4/Network-Analysis-I/master/Data/polblogs.txt", sep = " ", header = TRUE, stringsAsFactors = FALSE)
pblog.edgelist <- as.matrix(pblog.data) + 1

pblog.labels <- as.matrix(read.table("https://raw.githubusercontent.com/jdwilson4/Network-Analysis-I/master/Data/polblogs_labels.txt", header = FALSE, stringsAsFactors = FALSE))

# set colors 
# 0 --> 4 = blue
pblog.labels <- replace(pblog.labels, pblog.labels == 0, 4)
# 0 --> 2 = red
pblog.labels <- replace(pblog.labels, pblog.labels == 1, 2)


```

### Convert Network Edge Data
```{r, echo = TRUE, eval = TRUE}
# create igraph from edgelist
pblog.igraph <- graph.edgelist(pblog.edgelist)
# adjacency matrix from igraph
pblog.adjacency <- as_adj(pblog.igraph)

# create statnet object from edgelist
pblog.network <- network(pblog.edgelist)
plot(pblog.network, main = paste("Political Blog Network"), usearrows = TRUE, edge.col = "grey50", vertex.col = pblog.labels)

# remove isolates
connected.nodes <- degree(pblog.network) >= 1
sub.pblog.igraph <- induced.subgraph(pblog.igraph, V(pblog.igraph)[connected.nodes])
sub.pblog.edgelist <- as_edgelist(sub.pblog.igraph)
sub.pblog.adj <- as_adjacency_matrix(sub.pblog.igraph)

# convert to statnet network
sub.pblog.network <- network(sub.pblog.edgelist)

# subset labels of connected nodes
sub.labels <- subset(pblog.labels, degree(pblog.network) >= 1)

# plot subnetwork
sub.pblog.x <- plot(sub.pblog.network, main=paste("Political Blog Subnetwork"), usearrows=TRUE, edge.col = "grey50", vertex.col=sub.labels)
sub.pblog.x

```

---


##Network Centrality Measures and Visualization

###Calculate In-Degree, Out-Degree, Eigenvector, Betweenness, and Closeness Centralities
```{r, echo = TRUE, eval = TRUE}
#in degree centrality
in.degree.centrality <- colSums(as.matrix(sub.pblog.adj))

#out degree centrality
out.degree.centrality <- rowSums(as.matrix(sub.pblog.adj))

#eigenvector centrality
eigenvector.centrality <- eigen_centrality(sub.pblog.igraph, directed = FALSE)$vector

#normalized betweenness centrality
betweenness.centrality <- estimate_betweenness(sub.pblog.igraph, directed = TRUE, cutoff = 10)

#closeness centrality
closeness.centrality <- estimate_closeness(sub.pblog.igraph, mode = "total", normalized = FALSE, cutoff = 10)
```

###Plot Network According to Centrality Measures
```{r, echo = TRUE}
par(mfrow = c(2,3))

#plot original network 
plot(sub.pblog.network, main = paste("Political Blog Subnetwork"), usearrows = TRUE, edge.col = "grey50", coord = sub.pblog.x, vertex.col = sub.labels)

#plot in-degree
plot(sub.pblog.network, main = paste("In-Degree Centrality"), usearrows = TRUE, vertex.cex = in.degree.centrality / 50, edge.col = "grey50", coord = sub.pblog.x, vertex.col = sub.labels)

#plot out-degree
plot(sub.pblog.network, main = paste("Out-Degree Centrality"), usearrows = TRUE, vertex.cex = out.degree.centrality / 50, edge.col = "grey50", coord = sub.pblog.x, vertex.col = sub.labels)

#plot eigenvector
plot(sub.pblog.network, main = paste("Eigenvector Centrality"), usearrows = TRUE, vertex.cex = eigenvector.centrality*4 + 1, edge.col = "grey50", coord = sub.pblog.x, vertex.col = sub.labels)

#plot betweenness
plot(sub.pblog.network, main = paste("Betweenness Centrality"), usearrows = TRUE, vertex.cex = betweenness.centrality / 15000, edge.col = "grey50", coord = sub.pblog.x, vertex.col = sub.labels)

#plot closeness
plot(sub.pblog.network, main = paste("Closeness Centrality"), usearrows = TRUE, vertex.cex = closeness.centrality * 150 + 1, edge.col = "grey50", coord = sub.pblog.x, vertex.col = sub.labels)

```